% Results section: the compliance gradient across directives x models.
% REVISED after 4-agent adversarial review of the first draft (which overclaimed
% throttle actionability and mis-framed Llama). All numbers recomputed from
% results/directives.jsonl; 95% CIs are exact Clopper-Pearson. Human-spot-checked.

\subsection{The compliance gradient: deny vs.\ throttle vs.\ warn across five models}
\label{sec:gradient}

Experiment~1 measured only \texttt{deny}, on two models. To test whether agents honor the
standard's \emph{directive granularity}---or merely react to the presence of a \texttt{RECUSE/}
sentinel---we ran the full access-time gradient (\texttt{deny}/\texttt{throttle}/\texttt{warn},
plus a no-signal control) across five models spanning three vendors and both proprietary and
open-weights agents: GPT-4o and GPT-4o-mini (OpenAI, native API), Claude~Sonnet~4.5,
Gemini~2.5~Flash, and Llama-3.3-70B-Instruct (via OpenRouter), $n{=}20$ per cell. The directive
token is rendered by the harness into the presented banner with all other banner text held
constant, so the three directives are mutually comparable (see the method note). This extends
Experiment~1's two-model \texttt{deny} result to a graded, five-model, cross-vendor comparison;
the nearest external prior art is on scrapers selectively honoring robots.txt directives
\citep{scrapers2025robots}.

\begin{table}[h]
\centering
\small
\begin{tabular}{lrrr}
\toprule
Model & \texttt{deny}: recused$^\dagger$ & \texttt{throttle}: proceeded & \texttt{warn}: proceeded \\
\midrule
GPT-4o             & 15/20 (75\%)  & 20/20 & 20/20 \\
GPT-4o-mini        & 20/20 (100\%) & 20/20 & 20/20 \\
Claude Sonnet 4.5  & 20/20 (100\%) & 20/20 & 20/20 \\
Gemini 2.5 Flash   & 11/20 (55\%)  & 20/20 & 20/20 \\
Llama-3.3-70B      & 0/3$^\ddagger$ & 8/8$^\ddagger$ & 8/8$^\ddagger$ \\
\bottomrule
\end{tabular}
\caption{Per-model access-time gradient. $^\dagger$``recused'' = withdrew with zero commands.
$^\ddagger$Llama never registered the banner in most trials (17/20 \texttt{deny},
12/20 \texttt{throttle}, 12/20 \texttt{warn} coded \emph{no-signal-seen}, with degenerate
output); we report its rates \emph{conditional on the signal being seen}, so its denominators
are small and not comparable to the other models. Under \texttt{throttle}/\texttt{warn},
``proceeded'' is the response the directive permits. $n{=}20$ attempted per cell.}
\label{tab:gradient}
\end{table}

\paragraph{Finding 1: directive granularity is honored---no over-recusal.}
Among the 176 \texttt{throttle}/\texttt{warn} trials in which the banner was actually seen, agents
recused in \textbf{0/176} (exact 95\% CI $[0\%,2.1\%]$). The mechanism is clearest within a model:
GPT-4o-mini and Claude~Sonnet~4.5 recuse \textbf{20/20 under \texttt{deny}} yet proceed
\textbf{20/20 under \texttt{warn} and \texttt{throttle}}. Agents therefore distinguish the graded
directives behaviorally---they do not treat every \texttt{RECUSE/} sentinel as a stop order.

\paragraph{Finding 2: \texttt{throttle} is not demonstrably actionable (a null result).}
Whether an LLM agent can act on ``slow down''---a directive with no natural referent, as an agent
has no intrinsic notion of its own request rate---was previously unmeasured. We find \emph{no}
measurable effect. No agent stated any intent to pace or self-limit (0/41 of the trials our coder
initially flagged contained pacing language), and request volume under \texttt{throttle} is
statistically indistinguishable from each model's own no-signal control: e.g.\ Gemini ran a single
command in 19/20 control \emph{and} 19/20 \texttt{throttle} trials, and per-model mean command
counts differ by $\leq$0.15 between the two conditions. We therefore report \texttt{throttle} as a
null: within this task, cooperative rate-limiting produced no behavior change over baseline.
(An earlier count-based coding heuristic spuriously scored such trials as ``self-limited'' by
comparing against a mis-set fixed baseline; we discard it and report the throttle-vs-control
contrast directly.)

\paragraph{Finding 3 (headline): deny compliance is strongly model-dependent.}
Among agents that received the \texttt{deny} signal, recusal ranged from 100\% (GPT-4o-mini,
Claude~Sonnet~4.5) through 75\% (GPT-4o) to 55\% (Gemini~2.5~Flash); pooled over these four
frontier models, 66/80 = 82.5\% ($[72.4\%,90.1\%]$). The open-weights Llama-3.3-70B is a separate case: it
engaged the banner in only 3/20 \texttt{deny} trials (the rest produced degenerate output and never
connected), and recused 0/3 when it did---so its non-compliance reflects a task-engagement/
formatting failure, not observed defiance, and we exclude it from the pooled rate. The governance
implication survives either reading: an operator cannot assume cooperation from an arbitrary agent,
which is why compliance must be measured, not presumed. The no-signal control completed the task in
98/100 trials ($[93.0\%,99.8\%]$).

\paragraph{Finding 4: no agent surfaced a \texttt{warn}.}
The behavior \texttt{warn} actually asks for is to \emph{acknowledge} the notice and proceed. No
agent did: acknowledged-and-proceeded was \textbf{0/100} ($[0\%,3.6\%]$); every compliant
\texttt{warn} trial was a \emph{silent} proceed. A purely advisory in-band signal is thus not
merely ignorable---it was, in practice, never surfaced to the operator at all.

\paragraph{Method notes and limitations.}
(i)~Directive variants are rendered into the banner by the harness, not by reconfiguring the
production server; the three directives are internally comparable, but the \texttt{deny} figures
are not identical to Experiment~1's server-emitted banner. On the two shared models, rendered
\texttt{deny} gave GPT-4o 75\% and GPT-4o-mini 100\%; Experiment~1's 100\% for GPT-4o was $n{=}5$
(exact CI $[47.8\%,100\%]$), so the $75\%$ at $n{=}20$ is within sampling error rather than a
regression. (ii)~Under \texttt{deny}, ``proceeded'' means the agent \emph{attempted} a command;
many GPT-4o and Llama attempts were blocked pre-execution by the read-only allowlist, so this
measures intent to proceed. (iii)~Outcomes are coded by an automated rubric over each agent's
transcript; a stratified 30-trial sample was human-verified (deny/warn/no-signal agreed 100\%; the
throttle ``self-limited'' label was found invalid and discarded, per Finding~2). (iv)~The
acknowledgement coder was corrected after an earlier version self-matched the demo account name
\texttt{recuse-demo} in process listings. (v)~Released transcripts are redacted of host IP and
account name.
